Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-22 docs/manuscript/manuscript.tex

& \\\bottomrule
  \end{tabular}
  \caption{Summary of water accessibility variables (national averages at household level)}
  \label{tab:summary}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{5cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{imws} & Households using an improved water source
 & Percentage of households whose main source of drinking water is an improved source.
 \\\hline
  \texttt{pipy} & Households using water piped into yard/plot
& Percentage of households whose main source of drinking water is water piped into the yard or plot.
\\\hline
 \texttt{truc} & Households using tanker truck & Percentage of households whose main source of drinking water is tanker truck.
\\\hline
 \texttt{ctan} & Households using cart with tank & Percentage of households whose main source of drinking water is cart with tank.
\\\hline
 \texttt{bott} & Households using bottled water & Percentage of households whose main source of drinking water is bottled water/demi john.
\\\hline
 \texttt{uiws} & Households using an unimproved water source & Percentage of households whose main source of drinking water is an unimproved source.
\\\hline
 \texttt{othw} & Households using other water source & Percentage of households whose main source of drinking water is other.
\\\hline
 \texttt{iwsp} & Households with improved water source on the premises & Percentage of the de jure population living in households with an improved water source on the premises.
\\\hline
 \texttt{basw} & Households with basic water service
 & Percentage of households with basic water service, defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
\\\hline
 \texttt{liws} & Households with limited water service
 & Percentage of households with limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
\\\hline
 \texttt{prem} & Households with water on the premises
 & Percentage of households with water on the premises.
\\\hline
 \texttt{bicy} & Households possessing a bicycle
 & Percentage of households possessing a bicycle.
\\\hline
 \texttt{acar} & Households possessing an animal drawn cart
 & Percentage of households possessing an animal drawn cart.
\\\hline
 \texttt{mcyc} & Households possessing a motorcycle
 & Percentage of households possessing a motorcycle.
\\\hline
 \texttt{pcar} & Households possessing a private car
 & Percentage of households possessing a private car.
\\\hline
 \texttt{boat} & Households possessing a boat with a motor
 & Percentage of households possessing a boat with a motor.
\\\hline
 \texttt{wigc} & Wealth index Gini coefficient
 & The Gini coefficient indicates the level of concentration of wealth in the country.
\\\hline
 & \\\bottomrule
  \end{tabular}
  \caption{Summary of explanatory variables (national averages measured as percentage at household level)}
  \label{tab:summary}
\end{table}
 \subsection*{Summary of variables}
\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{c c c}\toprule
    \bf Methods & \bf KL (clusters) & \bf Value Index \\\midrule
    \texttt{Ward.D2} & 4 & 2.4982 \\\hline
    \texttt{Ward.D} & 4 & 3.3397 \\\hline
    \texttt{Complete} & 4 & 3.9093 \\\hline
     \texttt{K-Means} & 12 & 8.6836  \\\bottomrule
  \end{tabular}
   \centering
  \begin{tabular}{c c c}\toprule
    \bf Methods & \bf Ratkowsky (clusters) & \bf Value Index \\\midrule
    \texttt{Ward.D2} & 4 & 0.2981 \\\hline
    \texttt{Ward.D} & 4 & 0.2981 \\\hline
    \texttt{Complete} & 3 & 0.3193 \\\hline
     \texttt{K-Means} & 3 & 0.2948 \\\bottomrule
  \end{tabular}
   \centering
  \begin{tabular}{c c c c c}\toprule
    \bf Methods & \bf Marriot(clusters) & \bf Value Index & \bf Gap & \bf Majority Rule  \\\midrule
    \texttt{Ward.D2} & 4 & 1.54E+40 &  8 & 4 \\\hline
    \texttt{Ward.D} & 4 & 1.79E+40 & 8 & 4 \\\hline
    \texttt{Complete} & 4 & 1.77E+40  & 7 & 4 \\\hline
     \texttt{K-Means} & 4 & 1.54E+40  & \texttt{N/A} & 3  \\\bottomrule
  \end{tabular}
  \caption{Summary of optimal clustering index values (Different Clustering Methods)}
  \label{tab:summary}
\end{table}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-time}
  \caption{Advanced Correlation Graph of Modes of Transportation and Time}
  \label{fig:transporation-time}
\end{figure}
Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01. Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources. Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001. The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources. 
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-1}
  \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
  \label{fig:transporation-water}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-2}
  \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
  \label{fig:transporation-water}
\end{figure}

@@

\subsection*{Water accessibility typologies}
\hl{Show the typology/cluster results and discuss.}
NBClust statistic analysis developed and determined four distinct groups. First cluster group composed of countries near the equator or in the Sub-Sahara Regions. The second analysis of the second group represents countries considered developed or recently industrialized.

%Results A gap statistic analysis resulted first in two distinct groups,with the first being composed of countries typically considered developed or newly industrialized.  The second analysis of the second group resulted in four country clusters. The greater subdivision would have resulted in statistically similar clusters, while greater aggregation would have resulted in highly internally heterogeneous clusters. The dendrogram resulting from this cluster analysis is shown in Fig. 1, and the corresponding gap curve yielding four clusters is shown in Fig. 2. Table 3 shows the resulting five water and sanitation country clusters. Fig. 3 shows the cluster country membership on a world map. The cluster indicator variable means are summarized in Table 4. The Dunn and Davies-BouldinIndices for our clusters, as well as the World Bank, WHO, and MDGgroupings are shown in Table 5. Our clusters have both the largest Dunn index and smallest Davies-Bouldin index, indicating that they represent a superior grouping along with the chosen relevant indicators.

>>>>

& \\\bottomrule
  \end{tabular}
  \caption{Summary of water accessibility variables (national averages at household level)}
  \label{tab:summary-wa}
\end{table}

\begin{table}[h!]
  \centering
  \begin{tabular}{c c m{5cm}}\toprule
    \bf Variable & \bf Description & \bf Definition \\\midrule
    \texttt{imws} & Households using an improved water source
 & Percentage of households whose main source of drinking water is an improved source.
 \\\hline
  \texttt{pipy} & Households using water piped into yard/plot
& Percentage of households whose main source of drinking water is water piped into the yard or plot.
\\\hline
 \texttt{truc} & Households using tanker truck & Percentage of households whose main source of drinking water is tanker truck.
\\\hline
 \texttt{ctan} & Households using cart with tank & Percentage of households whose main source of drinking water is cart with tank.
\\\hline
 \texttt{bott} & Households using bottled water & Percentage of households whose main source of drinking water is bottled water/demi john.
\\\hline
 \texttt{uiws} & Households using an unimproved water source & Percentage of households whose main source of drinking water is an unimproved source.
\\\hline
 \texttt{othw} & Households using other water source & Percentage of households whose main source of drinking water is other.
\\\hline
 \texttt{iwsp} & Households with improved water source on the premises & Percentage of the de jure population living in households with an improved water source on the premises.
\\\hline
 \texttt{basw} & Households with basic water service
 & Percentage of households with basic water service, defined as an improved water source with either water on the premises or round-trip collection time is 30 minutes or less.
\\\hline
 \texttt{liws} & Households with limited water service
 & Percentage of households with limited water service, defined as an improved water source with round-trip collection time greater than 30 minutes.
\\\hline
 \texttt{prem} & Households with water on the premises
 & Percentage of households with water on the premises.
\\\hline
 \texttt{bicy} & Households possessing a bicycle
 & Percentage of households possessing a bicycle.
\\\hline
 \texttt{acar} & Households possessing an animal drawn cart
 & Percentage of households possessing an animal drawn cart.
\\\hline
 \texttt{mcyc} & Households possessing a motorcycle
 & Percentage of households possessing a motorcycle.
\\\hline
 \texttt{pcar} & Households possessing a private car
 & Percentage of households possessing a private car.
\\\hline
 \texttt{boat} & Households possessing a boat with a motor
 & Percentage of households possessing a boat with a motor.
\\\hline
 \texttt{wigc} & Wealth index Gini coefficient
 & The Gini coefficient indicates the level of concentration of wealth in the country.
\\\hline
 & \\\bottomrule
  \end{tabular}
  \caption{Summary of explanatory variables (national averages measured as percentage at household level)}
  \label{tab:summary-var}
\end{table}
 \subsection*{Summary of variables}
\hl{Here, you show and discuss the data distributions (e.g. histograms, correlation plots, etc)}

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{c c c}\toprule
    \bf Methods & \bf KL (clusters) & \bf Value Index \\\midrule
    \texttt{Ward.D2} & 4 & 2.4982 \\\hline
    \texttt{Ward.D} & 4 & 3.3397 \\\hline
    \texttt{Complete} & 4 & 3.9093 \\\hline
     \texttt{K-Means} & 12 & 8.6836  \\\bottomrule
  \end{tabular}
   \centering
  \begin{tabular}{c c c}\toprule
    \bf Methods & \bf Ratkowsky (clusters) & \bf Value Index \\\midrule
    \texttt{Ward.D2} & 4 & 0.2981 \\\hline
    \texttt{Ward.D} & 4 & 0.2981 \\\hline
    \texttt{Complete} & 3 & 0.3193 \\\hline
     \texttt{K-Means} & 3 & 0.2948 \\\bottomrule
  \end{tabular}
   \centering
  \begin{tabular}{c c c c c}\toprule
    \bf Methods & \bf Marriot(clusters) & \bf Value Index & \bf Gap & \bf Majority Rule  \\\midrule
    \texttt{Ward.D2} & 4 & 1.54E+40 &  8 & 4 \\\hline
    \texttt{Ward.D} & 4 & 1.79E+40 & 8 & 4 \\\hline
    \texttt{Complete} & 4 & 1.77E+40  & 7 & 4 \\\hline
     \texttt{K-Means} & 4 & 1.54E+40  & \texttt{N/A} & 3  \\\bottomrule
  \end{tabular}
  \caption{Summary of optimal clustering index values (Different Clustering Methods)}
  \label{tab:summary-optimal}
\end{table}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-time}
  \caption{Advanced Correlation Graph of Modes of Transportation and Time}
  \label{fig:transporation-time}
\end{figure}
Figure 5 shows that Households possessing a motorcycle (mcyc) has a positive correlation with Households using a tube well/borehole (bore) and Households using rainwater (rain) with a p-value of less than 0.01. Similar to the previous analysis, there is a relationship between a mode of transportation to retrieve potential water sources. Figure 6 also shows that households possessing an animal-drawn cart (acar) have a positive and strong relationship with households using unprotected well water (uwel) with a p-value of less than 0.001. The results demonstrate that different modes of transportation factor into the accessibility of retrieving water sources on a global scale, therefore, there should be consideration of these variables to increase access to improved water sources. 
\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-1}
  \caption{Advanced Correlation Graph of Modes of Transportation and Water Resources}
  \label{fig:transporation-water-one}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{transportation-water-2}
  \caption{Advanced Correlation Graph of Modes of Transportation and Additional Water Resources}
  \label{fig:transporation-water-two}
\end{figure}

@@

\subsection*{Water accessibility typologies}
\hl{Show the typology/cluster results and discuss.}
NBClust statistic analysis developed and determined four distinct groups. First cluster group composed of countries near the equator or in the Sub-Sahara Regions. The analysis of the second group represents countries considered developed or recently industrialized. Third cluster group represents most of the countries in the central Africa and may have similar factors as the first cluster group. The fourth cluster group represents most of countries in the middle east or predominately non-African countries. There is a great internally heterogeneous clusters for the 76 distinct countries. The cluster analysis' dendrogram is shown in \autoref{fig:dendrogram}, and the corresponding NBClust curve yielding the optimal four clusters is shown in \autoref{fig:optimal-clusters, fig:optimal-cuts}.

%   Table 3 shows the resulting five water and sanitation country clusters. Fig. 3 shows the cluster country membership on a world map. The cluster indicator variable means are summarized in Table 4. The Dunn and Davies-BouldinIndices for our clusters, as well as the World Bank, WHO, and MDGgroupings are shown in Table 5. Our clusters have both the largest Dunn index and smallest Davies-Bouldin index, indicating that they represent a superior grouping along with the chosen relevant indicators.

\begin{figure}[h!]
  \centering
  \includegraphics[width=.7\textwidth]{optimal-clusters-ward-d2}
  \caption{Optimal Number of Clusters for Ward.D2}
  \label{fig:optimal-clusters}
\end{figure}
